### 2024-12-17 NLP/exercises.tex
Merged exercises and removed redundancy.

\begin{exercise}{Web as an application area for NLP}
  The web is an application area for NLP. Provide examples of applications.

  \begin{solution}
    Examples include:
    \begin{itemize}
        \item Internet of Services
        \item Community mining
        \item Information retrieval
        \item Machine translation for web-based content
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Web as a resource for NLP}
  How can the web improve the quality of NLP? Provide examples.

  \begin{solution}
    Examples include:
    \begin{itemize}
        \item Web as Corpus
        \item Analyzing web-based knowledge repositories
        \item Extracting structured information from unstructured web data
        \item Using web-scale data for training language models
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Morphological analysis}
  Identify all stems and affixes (prefix, suffix, infix, circumfix) in the following words: "index," "incorrect," "interesting."

  \begin{solution}
    Morphological breakdown:
    \begin{itemize}
        \item \textbf{index}: stem = "index"
        \item \textbf{incorrect}: prefix = "in," stem = "correct"
        \item \textbf{interesting}: stem = "interest," suffix = "ing"
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Stemming vs lemmatization}
  In contrast to lemmatization, stemming does not necessarily return a valid word form. Why is stemming still useful?

  \begin{solution}
    Stemming is still useful because:
    \begin{itemize}
        \item It is faster and computationally less expensive.
        \item It simplifies implementation for applications like Information Retrieval.
        \item It can still group similar words under a common base, which is often sufficient.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Corpus definition and usage}
  What is a corpus, and how can it be built and used for linguistic analysis?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Corpus definition}: A corpus is a large and structured set of texts, often used for linguistic research.
        \item \textbf{How to build}: Collect text data, preprocess it (cleaning, normalization), and annotate it for specific tasks.
        \item \textbf{Usage}: Analyze linguistic patterns, train NLP models, and evaluate machine learning algorithms.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{NLP datasets}
  What are NLP datasets, and why are they important?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Definition}: NLP datasets are collections of annotated or unannotated text data designed for specific NLP tasks (e.g., sentiment analysis, translation).
        \item \textbf{Importance}:
        \begin{itemize}
            \item Serve as a foundation for training and testing machine learning models.
            \item Enable reproducibility in research.
            \item Facilitate comparisons between models and techniques.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Corpus and dataset annotation}
  What is the annotation process for a corpus or dataset, and how does it impact machine learning applications?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Annotation process}: Adding labels or metadata to text data (e.g., part-of-speech tags, sentiment labels, named entities).
        \item \textbf{Impact on machine learning}:
        \begin{itemize}
            \item Improves model performance by providing high-quality training data.
            \item Enables supervised learning by associating input with output labels.
            \item Impacts scalability and generalization of models depending on annotation quality.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Large-scale dataset collection using crowdsourcing}
  What are the main strategies for large-scale dataset collection using crowdsourcing?

  \begin{solution}
    Strategies include:
    \begin{itemize}
        \item \textbf{Task design}: Clearly define annotation tasks with examples and guidelines.
        \item \textbf{Platform selection}: Use platforms like Amazon Mechanical Turk or Prolific.
        \item \textbf{Quality control}: Use redundancy (multiple annotators per task) and include test questions to ensure reliability.
        \item \textbf{Data aggregation}: Combine annotations using majority voting or statistical models.
    \end{itemize}
  \end{solution}
\end{exercise}

% Q: Why are more features not always better for learning a
% classifier?
% A: More training data might be needed; Dependent features
% might lead to deficient models; Some ML algorithms can only deal
% with specific data types; Slower.
% Q: What is the typical type of input data that a sequence tagger
% training step requires?
% A: Labeled training text. (e.g. POS, NER)

@@

\begin{exercise}{Tokenization}
  Give at least 3 examples of ambiguity in tokenization.

  \begin{solution}
    \begin{enumerate}
        \item \textbf{Hyphenated words}: Does "state-of-the-art" count as one token or three?
        \item \textbf{Contractions}: Should "don't" be treated as one token or split into "do" and "n't"?
        \item \textbf{Punctuation}: How should punctuation like periods in "U.S." or apostrophes in possessives ("John's book") be handled?
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Parse trees}
  "\textit{John shot the man with the gun}". Draw all possible parse trees for this sentence using the English grammar. What types of syntactic ambiguity are present?

@@

\begin{exercise}{Stemming and lemmatization}
  What is the difference between stemming and lemmatization?

  \begin{solution}
    While stemming reduces a word to its base form by removing pre- and suffices ("running" → "run"), lemmatization considers the word's context and grammar to find its dictionary form ("better" → "good"). Lemmatization is more precise than stemming but computationally more expensive.
  \end{solution}
\end{exercise}

\begin{exercise}{Web as corpus}
  List 4 issues you have to deal with when using the web as a corpus for NLP purposes and explain each.

  \begin{solution}
    \begin{enumerate}
        \item \textbf{Noise}: Content may contain errors, irrelevant information, or non-standard grammar.
        \item \textbf{Duplicates}: Identical or nearly identical data appears multiple times.
        \item \textbf{Bias}: Web content often reflects societal biases, which can skew analyses.
        \item \textbf{Reproducability}: Web pages change frequently, making reproducibility challenging.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Cross-language QA}
  How does cross-language QA work and what are the main issues?

  \begin{solution}
    Cross-language QA involves answering questions in one language based on information in another. This is accomplished by translating the question, retrieving relevant information in the target language, and generating an answer. Challenges include translation errors, cultural differences.
  \end{solution}
\end{exercise}

\begin{exercise}{Machine learning pipeline}
  What are the main concepts in a machine learning pipeline?

  \begin{solution}
    Data Preprocessing $\to$ Feature extraction $\to$ Model training $\to$ Evaluation $\to$ Deployment
  \end{solution}
\end{exercise}

\begin{exercise}{IR evaluation}
  What are the problems with precision, recall, and F-measure in IR? How can these be improved? Why is Cohen's Kappa better?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Precision}: How relevant are the retrieved documents? Favors retrieving fewer documents.
        \item \textbf{Recall}: How many of the relevant documents were retrieved? Favors retrieving more documents.
        \item \textbf{F-measure}: Harmonic mean of precision and recall. Balances the two metrics but not adjustable.
        \item \textbf{Cohen's Kappa}: Measures inter-annotator agreement, accounting for chance agreement. Better for evaluating IR systems because it considers agreement beyond chance.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{IR evaluation}
  Calculate P@5, P@10, and MAP for the given IR ranked result list. Analyze both IR system A and system B.

  \begin{solution}
    For P@5, P@10, and MAP:
    \begin{itemize}
        \item \textbf{P@5}: Count relevant documents in the top 5 results and divide by 5.
        \item \textbf{P@10}: Same calculation for the top 10 results.
        \item \textbf{MAP}: Average Precision for each query, then average across all queries.
    \end{itemize}
    Steps:
    \begin{itemize}
        \item Identify positions of relevant documents in ranked lists.
        \item Calculate precision at each relevant position.
        \item Take mean of these precision values.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Naive Bayes and SVM}
  Which of the following statements are correct?
  \begin{enumerate}
    \item Features for Naive Bayes classifiers do not need to be uncorrelated.
    \item Naive Bayes is slow to train.
    \item SVM itself is a multi-label classifier.
    \item SVM handles sparse data well.
  \end{enumerate}

  \begin{solution}
    Correct statements:
    \begin{enumerate}
        \item \textbf{True}: Features for Naive Bayes do not need to be uncorrelated, though it assumes independence for computation.
        \item \textbf{False}: Naive Bayes is fast to train, not slow.
        \item \textbf{False}: SVM is not inherently multi-label but can be adapted for multi-label classification.
        \item \textbf{True}: SVM handles sparse data well due to its kernel functions.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Corpus construction}
  How can a corpus be constructed for linguistic purposes from web data, and how can it be made usable?

  \begin{solution}
    Steps for constructing a corpus:
    \begin{enumerate}
        \item \textbf{Data collection}: Crawl web pages, scrape data, or gather domain-specific text.
        \item \textbf{Cleaning}: Remove duplicates, irrelevant content, and normalize text.
        \item \textbf{Annotation}: Add metadata or labels for tasks like classification or parsing.
        \item \textbf{Usability}: Store in accessible formats (e.g., JSON, XML) and document the corpus for ease of use.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Pragmatics and discourse analysis}
  Why are pragmatics and discourse analysis difficult?

  \begin{solution}
    Challenges in pragmatics and discourse analysis:
    \begin{itemize}
        \item \textbf{Context dependency}: Understanding depends heavily on prior sentences or external knowledge.
        \item \textbf{Ambiguity}: Words and phrases can have multiple meanings depending on the situation.
        \item \textbf{Social norms}: Interpretations vary based on cultural or societal factors.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Hate speech}
  Define hate speech and explain how to construct a hate speech classifier.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Definition}: Hate speech involves language that attacks or discriminates against individuals or groups based on attributes like race, religion, ethnicity, gender, or sexual orientation.
        \item \textbf{Constructing a hate speech classifier}:
        \begin{enumerate}
            \item Collect labeled datasets with examples of hate speech and non-hate speech.
            \item Preprocess text (e.g., tokenization, removing stop words).
            \item Extract features (e.g., bag-of-words, TF-IDF, embeddings).
            \item Train a supervised machine learning model (e.g., Naive Bayes, SVM, or deep learning models like LSTMs or transformers).
            \item Evaluate the classifier using metrics like precision, recall, and F1-score.
        \end{enumerate}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Counter hate speech}
  What is counter hate speech, and how can a system be constructed for it?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Definition}: Counter hate speech is content designed to reduce the impact of hate speech, often by promoting positive, empathetic, or factual responses.
        \item \textbf{Constructing a counter hate speech system}:
        \begin{enumerate}
            \item Collect datasets of effective counter hate responses.
            \item Train models to generate or retrieve appropriate responses based on the hate speech content.
            \item Use sentiment analysis and empathy detection to craft responses.
            \item Deploy the system in platforms like social media or forums, ensuring ethical considerations are addressed.
        \end{enumerate}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Crowdsourced annotation}
  What is the workflow for crowdsourced annotation?

  \begin{solution}
    Steps in a crowdsourced annotation workflow:
    \begin{enumerate}
        \item \textbf{Task design}: Define the annotation task clearly with examples and instructions.
        \item \textbf{Platform setup}: Use a crowdsourcing platform (e.g., Amazon Mechanical Turk, Prolific).
        \item \textbf{Quality control}: Include test questions and redundancy (multiple annotators per item) to ensure data reliability.
        \item \textbf{Data aggregation}: Aggregate annotations using majority voting or weighted schemes.
        \item \textbf{Evaluation}: Analyze annotation quality and address inconsistencies.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Web corpus}
  Why do duplicates exist on the web, and how can they be detected?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Duplicates}: Duplicates arise from mirrored websites, copied content, or redundant crawls.
        \item \textbf{Detection methods}:
        \begin{itemize}
            \item Hashing (e.g., MD5 or SHA): Compare hash values of documents.
            \item Similarity measures: Use cosine similarity or Jaccard index on n-grams.
            \item Structural analysis: Detect similarity in HTML structure.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{QA vs IR}
  What are the differences between Question Answering (QA) and Information Retrieval (IR)?

  \begin{solution}
    Differences between QA (Question Answering) and IR (Information Retrieval):
    \begin{itemize}
        \item \textbf{Goal}:
        \begin{itemize}
            \item QA aims to provide direct answers.
            \item IR retrieves documents containing relevant information.
        \end{itemize}
        \item \textbf{Processing}:
        \begin{itemize}
            \item QA requires deeper natural language understanding.
            \item IR focuses on keyword matching and ranking.
        \end{itemize}
        \item \textbf{Applications}:
        \begin{itemize}
            \item QA: Chatbots, customer support.
            \item IR: Search engines.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Cross-language QA}
  What are the applications of multi-language QA, and what difficulties arise?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Applications}: Multi-language QA systems are used in global customer support, cross-border e-commerce, and multi-language search engines.
        \item \textbf{Difficulties}:
        \begin{itemize}
            \item Translation errors can distort queries or results.
            \item Cultural and linguistic differences complicate understanding and generating answers.
            \item Lack of high-quality bilingual corpora for training.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Language detection}
  How does language detection work?

  \begin{solution}
    Approaches for language detection:
    \begin{enumerate}
        \item \textbf{N-gram analysis}: Compare n-grams of text against language-specific frequency profiles.
        \item \textbf{Machine learning models}: Use classifiers trained on labeled text for different languages.
        \item \textbf{Lexicon-based methods}: Match words in the text to dictionaries of different languages.
        \item \textbf{Hybrid approaches}: Combine multiple methods for higher accuracy.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Factual QA vs opinionated QA}
  What are the differences between factual QA and opinionated QA, and which is easier to evaluate?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Factual QA}: Deals with objective information (e.g., "What is the capital of France?").
        \item \textbf{Opinionated QA}: Involves subjective answers (e.g., "Is Paris a good travel destination?").
        \item \textbf{Evaluation}: Factual QA is easier to evaluate as answers can be verified against ground truth. Opinionated QA requires subjective or aggregated assessments.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{RAG and LangChain}
  What is Retrieval-Augmented Generation (RAG), and how does LangChain assist in implementing it?

@@

\begin{exercise}{Sentiment analysis}
  Compare sentiment analysis using supervised approaches and lexicon-based approaches.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Supervised learning}: Uses labeled data to train machine learning models.
        \begin{itemize}
            \item \textbf{Advantages}: High accuracy with large datasets.
            \item \textbf{Drawbacks}: Relies on annotated data and is domain-dependent.
        \end{itemize}
        \item \textbf{Lexicon-based approaches}: Use sentiment dictionaries to determine polarity.
        \begin{itemize}
            \item \textbf{Advantages}: Simplicity and no training requirement.
            \item \textbf{Drawbacks}: Limited contextual understanding and poor performance on complex text.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Edit distance}
  Calculate the Minimum Edit Distance for "TAC" vs. "CAT" and list all possible word outcomes.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Minimum Edit Distance for "TAC" and "CAT"}:
        \[
        \begin{array}{c|c|c|c|c}
            &  & C & A & T \\
            \hline
            & 0 & 1 & 2 & 3 \\
            T & 1 & 1 & 2 & 2 \\
            A & 2 & 2 & 1 & 2 \\
            C & 3 & 2 & 2 & 2 \\
        \end{array}
        \]
        \item \textbf{Word outcomes}: Possible outcomes from edits:
        \begin{enumerate}
            \item CAT
            \item TAT
            \item TAC
            \item CAC
        \end{enumerate}
    \end{itemize}
  \end{solution}
\end{exercise}

\setcounter{section}{2021}
\sheet{1st exam}

\begin{exercise}{Corpus construction}
  How would you construct a corpus for linguistic purposes and how can it be made usable?

  \begin{solution}
    \begin{enumerate}
        \item \textbf{Data collection}: Gather data from relevant sources (e.g., web crawling, domain-specific text).
        \item \textbf{Cleaning and preprocessing}: Remove duplicates, non-linguistic content, and normalize text (e.g., case normalization, tokenization).
        \item \textbf{Annotation}: Add metadata or linguistic labels (e.g., POS tags, syntactic trees) to the text.
        \item \textbf{Storage}: Store the corpus in an accessible format like JSON, XML, or relational databases.
        \item \textbf{Documentation}: Provide guidelines and metadata about the corpus for user understanding.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Pragmatics and discourse analysis}
  Why are pragmatics and discourse analysis difficult?

>>>>

\begin{exercise}{NLP and the Web}
  \begin{enumerate}
    \item The web is an application area for NLP. Provide examples of applications.
    \item How can the web improve the quality of NLP? Provide examples.
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item Information extraction, sentiment analysis, machine translation, question answering, etc.
      \item Web as a corpus, analyzing web-based knowledge repositories, extracting structured information, using web-scale data for training language models.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Datasets}
  \begin{enumerate}
    \item What datasets are used in NLP research, and why are they important?
    \item What is the annotation process for a corpus or dataset, and how does it impact machine learning applications?
    \item What are the main strategies for large-scale dataset collection using crowdsourcing?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item NLP datasets are collections of annotated or unannotated text data designed for specific NLP tasks (e.g., sentiment analysis, translation). They serve as a foundation for training and testing machine learning models, enable reproducibility in research, and facilitate comparisons between models and techniques.
      \item Annotation process involves adding labels or metadata to text data (e.g., part-of-speech tags, sentiment labels). It improves model performance by providing high-quality training data, enables supervised learning by associating input with output labels, and impacts scalability and generalization of models depending on annotation quality.
      \item Strategies include task design (clearly define annotation tasks with examples and guidelines), platform selection (use platforms like Amazon Mechanical Turk or Prolific), quality control (use redundancy and include test questions to ensure reliability), and data aggregation (combine annotations using majority voting or statistical models).
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Text preprocessing}
  \begin{enumerate}
    \item Why are more features not always better for learning a classifier?
    \item What is the typical type of input data that a sequence tagger training step requires?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item More features can lead to overfitting, require more training data, introduce noise, and slow down training and inference.
      \item Labeled training text (e.g., part-of-speech tags, named entity recognition).
    \end{enumerate}
  \end{solution}
\end{exercise}

@@

\begin{exercise}{Tokenization ambiguity}\label{ex:tokenization}
  List three types of tokenization ambiguity and explain each.

  \begin{solution}
    \begin{itemize}
      \item \textbf{Hyphenation}: Is "state-of-the-art" one token or three?
      \item \textbf{Punctuation}: Should periods in abbreviations (e.g., "U.S.") be separated?
      \item \textbf{Contractions}: Should "don't" be split into "do" and "n't"?
      \item \textbf{Language differences}: Tokenization rules vary by language (e.g., Chinese characters vs. English words).
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Parsing ambiguity}
  \begin{quote}\center "\textit{John shot the man with the gun}"\end{quote}
  Draw all possible parse trees for this sentence using the English grammar. What types of syntactic ambiguity are present?

@@

\begin{exercise}{Stemming v. Lemmatization}\label{ex:stem-v-lemm}
  \begin{enumerate}
    \item What is the difference between stemming and lemmatization?
    \item In contrast to lemmatization, stemming does not necessarily return a valid word form. Why is stemming still useful?
    \item Identify all stems and affixes (prefix, suffix, infix, circumfix) in the following words: "index," "incorrect," "interesting."
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
        \item Stemming reduces words to their root form by removing affixes. Lemmatization finds the base or dictionary form of a word based on its context.
        \item Stemming is useful for information retrieval tasks where speed is crucial, and minor inaccuracies are acceptable.
        \item \begin{itemize}
            \item \textbf{index}: stem = "index"
            \item \textbf{incorrect}: prefix = "in," stem = "correct"
            \item \textbf{interesting}: stem = "interest," suffix = "ing"
        \end{itemize}
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Machine learning}\label{ex:ml}
  \begin{enumerate}
    \item Which of the following statements are correct? Give a brief explanation for each.
      \begin{itemize}
        \item[$\square$] Features for Naive Bayes classifiers do not need to be uncorrelated.
        \item[$\square$] Naive Bayes is slow to train.
        \item[$\square$] SVM itself is a multi-label classifier.
        \item[$\square$] SVM handles sparse data well.
      \end{itemize}
      \item Explain Support Vector Machines when you have more than two classes.
      \item What are the main stages in a machine learning pipeline?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item \begin{itemize}
          \item \textbf{True}: Features for Naive Bayes do not need to be uncorrelated, though it assumes independence for computation.
          \item \textbf{False}: Naive Bayes is fast to train, not slow.
          \item \textbf{False}: SVM is not inherently multi-label but can be adapted for multi-label classification.
          \item \textbf{True}: SVM handles sparse data well due to its kernel functions.
        \end{itemize}
      \item \begin{itemize}
          \item \textbf{One-vs-One (OvO)}: Trains a classifier for every pair of classes. Final classification is determined by majority voting.
          \item \textbf{One-vs-All (OvA)}: Trains one classifier per class against all other classes. The highest-confidence prediction is chosen.
        \end{itemize}
      \item Data Preprocessing $\to$ Feature extraction $\to$ Model training $\to$ Evaluation $\to$ Deployment
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Information retrieval}\label{ex:ir}
  Information retrieval (IR) is concerned with finding relevant information in large collections of data.
  \begin{enumerate}
    \item Describe stages in the IR cycle.
    \item Where is classification used in IR?
    \item What are key differences between Question Answering (QA) and IR?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
        \item \begin{itemize}
            \item \textbf{Indexing}: Process and store data for efficient retrieval.
            \item \textbf{Querying}: Users submit queries to search the index.
            \item \textbf{Retrieval}: Return ranked results based on relevance scores.
            \item \textbf{Evaluation}: Assess performance using metrics like precision and recall.
          \end{itemize}
        \item In query classification, document classification, and relevance classification.
        \item QA provides direct answers, requires natural language understanding, and is used in chatbots. IR retrieves documents, focuses on keyword matching, and powers search engines.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Corpus construction}\label{ex:corpus}
  \begin{enumerate}
    \item Define a corpus in the context of NLP.
    \item How would you construct a corpus usable for linguistic purposes?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item A corpus is a large collection of text data used for linguistic analysis, machine learning, or NLP research.
      \item \begin{enumerate}
          \item \textbf{Data collection}: Gather data from relevant sources (e.g., web crawling, domain-specific text).
          \item \textbf{Cleaning and preprocessing}: Remove duplicates, non-linguistic content, and normalize text (e.g., case normalization, tokenization).
          \item \textbf{Annotation}: Add metadata or linguistic labels (e.g., POS tags, syntactic trees) to the text.
          \item \textbf{Storage}: Store the corpus in an accessible format like JSON, XML, or relational databases.
          \item \textbf{Documentation}: Provide guidelines and metadata about the corpus for user understanding.
        \end{enumerate}
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Pragmatics and discourse analysis}\label{ex:pragmatics}
  Why are pragmatics and discourse analysis difficult?

  \begin{solution}
    \begin{itemize}
      \item \textbf{Context dependency}: Understanding depends heavily on prior sentences or external knowledge.
      \item \textbf{Ambiguity}: Words and phrases can have multiple meanings depending on the situation.
      \item \textbf{Social norms}: Interpretations vary based on cultural or societal factors.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Hate speech}\label{ex:hate-speech}
  \begin{enumerate}
    \item Define hate speech.
    \item How would you construct a hate speech classifier?
    \item Explain how to counter hate speech.
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
        \item Language that attacks or discriminates against individuals or groups based on attributes.
        \item Collect labeled datasets, preprocess text, extract features, train a model, and evaluate it.
        \item Collect examples of effective counter hate responses, classify hate speech, and generate counter-responses using NLP techniques.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Crowdsourcing}\label{ex:crowdsourcing}
  \begin{enumerate}
    \item What is the workflow for crowdsourced annotation?
    \item What are opportunities and challenges in using crowdsourcing for data annotation?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
        \item \textbf{Task design}: Define the annotation task clearly with examples and instructions.
        \item \textbf{Platform setup}: Use a crowdsourcing platform (e.g., Amazon Mechanical Turk, Prolific).
        \item \textbf{Quality control}: Include test questions and redundancy (multiple annotators per item) to ensure data reliability.
        \item \textbf{Data aggregation}: Aggregate annotations using majority voting or weighted schemes.
        \item \textbf{Evaluation}: Analyze annotation quality and address inconsistencies.
    \end{enumerate}
    \begin{itemize}
        \item[+] Fast and scalable data annotation.
        \item[+] Access to diverse annotators.
        \item[+] Cost-effective compared to hiring experts.
        \item[-] Quality control is challenging.
        \item[-] May introduce noise or bias.
        \item[-] Difficult to ensure annotator expertise.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Web as corpus}\label{ex:web-corpus}
  In NLP research, the web is often used as a corpus.
  \begin{enumerate}
    \item What are some opportunities of using the web as a corpus?
    \item The web contains many duplicates. How can they be detected?
    \item Explain other key challenges in using the web as a corpus.
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item Large-scale, diverse data sources; up-to-date content; covers multiple languages and domains.
      \item Hashing (e.g., MD5, SHA) to compare file contents; cosine similarity or Jaccard index for text comparisons; comparing HTML structure or metadata.
      \item \begin{itemize}
          \item \textbf{Noise}: Content may contain errors, irrelevant information, or non-standard grammar.
          \item \textbf{Duplicates}: Identical or nearly identical data appears multiple times.
          \item \textbf{Bias}: Web content often reflects societal biases, which can skew analyses.
          \item \textbf{Reproducibility}: Web pages change frequently, making reproducibility challenging.
        \end{itemize}
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Classification metrics}\label{ex:class-metrics}
  Have a look at the classification results in Table \ref{tbl:results}. \input{res/results.tbl}

  \begin{enumerate}
    \item Calculate Precision, Recall, and F1-Score. What is the problem with these metrics in IR?
    \item Calculate Cohen's Kappa. Why is it better for evaluating IR systems?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item \begin{itemize}
        \item \textbf{Precision} $\frac{TP}{TP + FP} = \frac{10}{10+5} = .6666$
        \item \textbf{Recall}: $\frac{TP}{TP + FN} = \frac{10}{10+3} = .7692$
        \item \textbf{F1-Score}: $2 \times \frac{\text{Precision} \times \text{Recall}}{\text{Precision} + \text{Recall}} = 2 \times \frac{.6666 \times .7692}{.6666 + .7692} = .7142$
        \end{itemize}
        These metrics are limited as they do not account for chance agreement.
      \item Cohen's Kappa is calculated as $\kappa = \frac{P_o - P_e}{1 - P_e}$ and describes the observed agreement $P_o$ relative to the expected agreement $P_e$.
        \[ \kappa = \frac{.7333-.5}{1-.5} = \frac{0.2333}{0.5} = 0.4666 \]
        This metric is better for evaluating IR systems because it considers agreement beyond chance.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Information retrieval metrics}\label{ex:ir-metrics}
  Two search engines have produced the following results ($+$ = relevant, $-$ = irrelevant).
  \begin{itemize}
    \item \textbf{Engine A}: $+, -, +, -, +, +, -, +, +, +$
    \item \textbf{Engine B}: $+, +, -, +, +, -, +, +, -, +$
  \end{itemize}

  \begin{enumerate}
    \item Calculate P@5, P@10 and MAP for both engines.
    \item How do these metrics help evaluate the performance of IR systems?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item \begin{itemize}
          \item \textbf{Engine A}:
          \begin{itemize}
            \item $P@5 = \frac{3}{5} = 0.6$
            \item $P@10 = \frac{7}{10} = 0.7$
            \item $MAP = \frac{1}{3} \left( \frac{1}{1} + \frac{2}{4} + \frac{3}{5} \right) = 0.5666$
          \end{itemize}
          \item \textbf{Engine B}:
          \begin{itemize}
            \item $P@5 = \frac{4}{5} = 0.8$
            \item $P@10 = \frac{7}{10} = 0.7$
            \item $MAP = \frac{1}{3} \left( \frac{1}{1} + \frac{2}{4} + \frac{3}{7} \right) = 0.5476$
          \end{itemize}
        \end{itemize}
      \item These metrics help evaluate IR systems by measuring precision at different cutoff points and the average precision across all relevant documents.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Language and plagiarism detection}\label{ex:detect}
  \begin{enumerate}
    \item What are the main approaches for language detection?
    \item And for plagiarism detection?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item \begin{itemize}
          \item \textbf{N-gram analysis}: Compare the frequency of character n-grams against language-specific profiles.
          \item \textbf{Machine learning models}: Train classifiers (e.g., SVM, neural networks) using labeled text data for various languages.
          \item \textbf{Lexicon-based methods}: Use dictionaries to match input text to language-specific words.
          \item \textbf{Hybrid approaches}: Combine n-gram, lexicon, and machine learning techniques for better accuracy.
        \end{itemize}
      \item \begin{itemize}
          \item \textbf{String matching}: Use algorithms to find exact matches between documents.
          \item \textbf{Semantic similarity}: Apply embeddings to detect paraphrased or rephrased content.
          \item \textbf{Citation analysis}: Check for missing or incorrect citations.
          \item \textbf{Fingerprinting}: Generate hashes of text fragments and compare across sources.
        \end{itemize}
      \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Question answering}\label{ex:qa}
  Question Answering (QA) systems aim to provide direct answers to user queries.
  \begin{enumerate}
    \item What are the differences between factual QA and opinionated QA? and which is easier to evaluate?
    \item Briefly explain QALD. What are key challenges?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item \begin{itemize}
          \item \textbf{Factual QA}: Answers objective questions with verifiable information.
          \item \textbf{Opinionated QA}: Provides subjective answers based on opinions or preferences.
          \item Factual QA is easier to evaluate as answers can be verified against ground truth.
        \end{itemize}
      \item QALD (Question Answering over Linked Data) focuses on answering questions using linked data (e.g., RDF datasets). Challenges include mapping natural language queries to structured data, handling incomplete or ambiguous data, and scaling to large and heterogeneous sources.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Cross-language QA}\label{ex:cross-lang-qa}
  Cross-language Question Answering (QA) is a challenging task.
  \begin{enumerate}
    \item Explain the process of cross-language question-answering.
    \item Where is it applied?
    \item Explain the challenges faced in cross-language QA.
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
        \item Translate the question, retrieve relevant information in the target language, generate an answer, and translate it back.
        \item Multi-language QA systems are used in global customer support, cross-border e-commerce, and multi-language search engines.
        \item Challenges include translation errors, lack of parallel resources, cultural differences, and mismatched linguistic structures.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{RAG and LangChain}\label{ex:rag}
  What is Retrieval-Augmented Generation (RAG), and how does LangChain assist in implementing it?

@@

\begin{exercise}{Sentiment analysis}\label{ex:sent-analysis}
  Sentiment analysis aims to determine the sentiment expressed in text.
  \begin{enumerate}
    \item Describe three applications of sentiment analysis.
    \item How do factual and opinionated sentiment analysis differ?
    \item Compare sentiment analysis using supervised approaches and lexicon-based approaches.
    \item How can lexicon-based sentiment analysis be improved using distributed thesauruses?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
        \item \begin{itemize}
            \item \textbf{Customer feedback}: Analyze reviews for product improvements.
            \item \textbf{Social media monitoring}: Detect trends and public sentiment.
            \item \textbf{Political analysis}: Gauge public opinion on policies or events.
          \end{itemize}
        \item \begin{itemize}
            \item \textbf{Factual sentiment analysis}: Focuses on extracting objective sentiment (e.g., product ratings).
            \item \textbf{Opinionated analysis}: Identifies subjective opinions (e.g., customer reviews).
          \end{itemize}
        \item \begin{itemize}
            \item \textbf{Supervised learning}: Uses labeled data to train machine learning models.
            \item \textbf{Lexicon-based approaches}: Use sentiment dictionaries to determine polarity.
          \end{itemize}
        \item Extend lexicons by identifying synonyms, antonyms, or similar words in distributed word embeddings like Word2Vec or GloVe.
      \end{enumerate}
  \end{solution}
\end{exercise}
